---
title: "Fricatives"
subtitle: ""
author: "Isabelle Chang"
institute: "Rutgers University"
date: "Last Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library("tidyverse")
library("here")
library("ds4ling")
library("patchwork")
library("untidydata")
library("broom")
library("emmeans")
```

```{r load_data, include=FALSE}
#load data from csv
raw_data <- read_csv(
  file=here("data_raw", "fricatives.csv"))

tidy_data <- raw_data
tidy_data %>%
  #fric: s, esh
  #names to = fric, measurement
  pivot_longer(cols=c("s_cog", "sh_cog", "s_skewness", "sh_skewness"),
               names_to = c("fric", "m_type"),
               names_sep = "_",
               values_to = "m_val") %>%
  arrange(obs, fric) %>%
  write_csv(file=here("data_tidy", "fricatives_tidy.csv"))
```

# Descriptive Statistics
```{r, message=FALSE, echo=FALSE}
#get descriptive stats
fric_stats <- read_csv(file=here("data_tidy", "fricatives_tidy.csv"))
fric_stats %>%
  group_by(fric) %>%
  summarize(
    avg_skew=mean(subset(m_val,m_type=="skewness")),
    sd_skew=sd(subset(m_val, m_type=="skewness")),
    avg_cog=mean(subset(m_val, m_type=="cog")),
    sd_cog=sd(subset(m_val, m_type=="cog"))) %>%
  write_csv(file=here("data_tidy", "fricatives_summary.csv"))
#put it in a table
fric_table <- read_csv(file=here("data_tidy", "fricatives_summary.csv"))
fric_table %>%
  mutate_if(is.numeric, round, digits=3) %>%
  arrange(fric) %>%
  knitr::kable()
```
---
# Center of Gravity as a Function of Phoneme

```{r, message=FALSE, echo=FALSE}
#x: s and sh
#y: the cog values
cog_plot <- read_csv(file=here("data_tidy", "fricatives_tidy.csv"))
cog_plot %>%
  #first make it smaller by grabbing subset of m_val
  subset(m_type == "cog") %>%
  #then plot
  ggplot() + 
  aes(x=factor(fric), 
      y=m_val,
      fill=factor(fric)) +
  xlab("Phoneme") +
  ylab("Center of Gravity (Hz)") +
  labs(fill="Phoneme") +
  geom_boxplot() +
  geom_point()
```

---

# Skewness as a Function of Phoneme

```{r, message=FALSE, echo=FALSE}
#x: s and sh
#y: the skewness values
skew_plot <- read_csv(file=here("data_tidy", "fricatives_tidy.csv"))
skew_plot %>%
  #first make it smaller by grabbing subset of m_val
  subset(m_type == "skewness") %>%
  #then plot, but with stats summary this time
  ggplot() + 
  aes(x=factor(fric), 
      y=m_val,
      fill=factor(fric)) +
  xlab("Phoneme") +
  ylab("Skewness") +
  labs(fill="Phoneme") +
  stat_summary(fun.data=mean_se, geom='pointrange', pch=21)
```

---
# Center of Gravity as a Function of Skewness for [s]- Model Summary
```{r message=FALSE, echo=FALSE}
#read in data
cog_skew <- read_csv(file=here("data_tidy", "fricatives_tidy.csv")) %>%
  subset(fric=="s") %>%
  pivot_wider(names_from="m_type", values_from="m_val")
#model
cogskew_mod <- lm(cog ~ 1 + skewness, data=cog_skew)
#table
tidy(cogskew_mod) %>%
  knitr::kable()
```
---
# Center of Gravity as a Function of Skewness for [s]- Scatter Plot
```{r message=FALSE, echo=FALSE, out.height="50%", out.width="50%"}
cogskew_plot <- cog_skew %>%
#scatter plot
  ggplot() +
  aes(x=round(skewness,2), y=round(cog,2)) +
  xlab("Skewness") +
  ylab("Center of Gravity (Hz)") +
  geom_point()

cogskew_plot
```
---
# Model Diagnostics
---
# Results
---
# Assumptions Dataset
```{r message=FALSE, echo=FALSE}
assumptions_plot <- read_csv(file=here("data_raw", "assumptions.csv")) %>%
  ggplot() +
  aes(x=x, y=y) +
  xlab("x") +
  ylab("y") +
  geom_point()

assumptions_plot
```

---

# Assumptions Dataset - Analysis

It would not be appropriate to fit a linear model to this dataset because it violates multiple assumptions we have to make when fitting a linear model:
- Visually, the relationship between each $x_i$ and $y_i$ value is nonlinear
- The model summary for a linear model fitted to $x_i$ and $y_i$ values shows that the mean of its residuals is non-zero.
```{r, message=FALSE, echo=FALSE, include=FALSE}
assumptions_data = read_csv(file=here("data_raw", "assumptions.csv"))

mod_assumptions <- lm(x~y, data=assumptions_data)

summary(mod_assumptions)
                           
```